\subsubsection{Conservation}
To assess the AAM conservation properties of the CAM-SE dynamical core, the diagnostics used in \cite{LBDL2014JAMES} are applied to the Aqua-planet simulation described in Section \ref{sec:APE}. In the dynamical core the column integrated wind and mass AAM are written to history tapes and the total integrals of AAM are computed as a post-processing step. The torques are obtained by subtracting the global integrals of AAM divided by the time-increment between the outputs of the AAM. For details on the discretized AAM diagnostics see \cite{LBDL2014JAMES}. These diagnostic outputs are part of the CESM2 release and are controlled with namelist variables.

As discussed in Section \ref{sec:aam} the total AAM torque from the dynamical core, in the absence of topography, should be small compared to the torque from the parameterizations. Figure \ref{fig:aam} shows that that is indeed the case for CAM-SE. The spurious torques from the dynamical core are approximately a factor 100 smaller than the physical torques from the parameterizations. These results are very similar to the dynamical core torques found in the dry Held-Suarez setup used in \citet{LBDL2014JAMES}.

Similarly to the AAM diagnostics, the total column-integrated moist energy is output at various places in the dynamical core and physics parameterizations. The dynamical core conserves the moist total energy, equation \eqref{eq:comprehensice_energy}, to about 0.1$W/m^2$. The frictional heating term described in section \ref{sec:frictional_heating} is approximately 0.4$W/m^2$ and hence an important term for total moist energy conservation. As mentioned in section \ref{sec:aam} the CAM physics energy fixer enforces a different energy than the comprehensive moist energy. This inconsistency should be removed in future CAM versions, however this is not a trivial modification to the CAM physics package. The discrepancy between the two definitions of energy is approximately 0.5$W/m^2$ \citep[similar to what ][ found when just including the correct heat capacity for water vapor in the total energy equation]{T2011LNCSEb}. A detailed energy analysis will be the subject of a future publication.
